---
title: "Activity monitoring data"
author: "Karnam Vasudeva Rao"
date: "February 7, 2016"
output: html_document
---

#Report
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.The variables included in this dataset are: steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format; interval: Identifier for the 5-minute interval in which measurement was taken. The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.





*Set working directory*

```{r}
setwd("C:/Vasu/R_Coursera/ReproducibleResearch/repdata-data-activity")
# list.files()
```


####1. Read the file

```{r}
data <- read.csv("activity.csv")
# head(data)
```


####2. What is mean total number of steps taken per day?
Histogram for total number of steps taken each day

```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)
#total.steps
qplot(total.steps, colour = I("blue"))+ 
  labs(title = "Total steps per day", binwidth = 500, x = "Steps per day", y = "Frequency")

```

####3.Mean and median total number of steps taken per day

```{r}
mean(total.steps, na.rm = TRUE)
# 9354.23
median(total.steps, na.rm = TRUE)
# 10395
```


####4. What is the average daily activity pattern?
```{r}
daily.average <- aggregate(x = list(steps = data$steps), by = list(interval = data$interval),FUN = mean, na.rm = TRUE)
head(daily.average)
```
####4. Time series plot of the average number of steps taken

```{r}
ggplot()+
  geom_line(data = daily.average, aes(x = interval, y = steps), colour = 'blue') +
  ggtitle("Average steps across all days") + 
  xlab("5-minute interval") +  
  ylab("Average steps taken") 

```


####5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
daily.average[which.max(daily.average$steps),]
# interval    steps
# 104      835 206.1698
```

####6. MISSING VALUES
*total number of missing values in the dataset* 

```{r}
missing.values <- length(which(is.na(data$steps)))
missing.values
#2304 rows have missing values
```
*Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated.For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.*

*Replace each missing value with the mean value of its 5-minute interval. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
fill <- function(steps, interval) {
  filled <- NA
  if (!is.na(steps))
    filled <- c(steps)
  else
    filled <- (daily.average[daily.average$interval==interval, "steps"])
  return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill, filled.data$steps, filled.data$interval)

```
####7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
steps.filled <- tapply(filled.data$steps, filled.data$date, sum)
qplot(steps.filled, xlab = 'Total steps per day (filled)', ylab = 'Frequency', binwidth = 500)
```

*mean and median total number of steps taken per day*
```{r}
steps.filledMean <- mean(steps.filled)
steps.filledMedian <- median(steps.filled)
```
*Are there differences in activity patterns between weekdays and weekends?*
*Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.*

```{r}
filled.data$datetype <-  ifelse(as.POSIXlt(filled.data$date)$wday %in% c(0,6), 'weekend', 'weekday')
```
####8. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 


```{r}
average.filled.data <- aggregate(steps ~ interval + datetype, data = filled.data, mean)
ggplot(average.filled.data, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(datetype ~ .) +
  xlab("5-minute interval") + 
  ylab("avarage number of steps")
```

